### 2021-06-04 paper/main.tex
added LaTeX file

\documentclass{article}

% Formatting
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage[titletoc,title]{appendix}
\usepackage{amsmath,amsfonts,amssymb,mathtools}
\usepackage{amsthm}
% \usepackage{minted}
\usepackage{polynom}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{prop}[theorem]{Proposition}

% Title content
\title{Development and Implementation of\\Novel Pair Value Encryption System}
\author{Rushil Mallarapu}
\date{June 6, 2021}

\begin{document}

\maketitle

\section{}

Use Euler's theorem to compute:

\begin{enumerate}
    \item $5^{60} \pmod{21}$
    \item $2^{35} \pmod{35}$
\end{enumerate}

\subsection{}

First, we verify that 5 and 21 are coprime, which they are. Then, we compute $\phi(21)$ via the standard route of decomposing 21 as its prime factorization $21 = 3\cdot 7$. Thus, $\phi(21) = \phi(3)\phi(7) = 2 \cdot 6 = 12$. Now, by Euler's theorem, it holds that $5^{12} \equiv 1 \pmod{21}$. Dividing 60 by 12 gives $60 = 5 \cdot 12 + 0$, by the division algorithm. Therefore, $5^{60} \equiv (5^{12})^5 \equiv 1^5 \equiv 1 \pmod{21}$. Thus, $5^{60} \equiv 1 \pmod{21}$.

\subsection{}

As before, we verify that 2 and 35 are coprime, which they are. Next, we compute $\phi(35)$ by decomposing 35 as its prime factorization $35 = 5 \cdot 7$. Thus, $\phi(35) = \phi(5)\phi(7) = 4 \cdot 6 = 24$. Now, by Euler's theorem, it holds that $2^{24} \equiv 1 \pmod{35}$. Dividing 35 by 24 gives $35 = 1 \cdot 24 + 11$, by the division algorithm. Therefore, $2^{35} \equiv (2^{24})^1 \cdot 2^{11} \equiv 1^1 \cdot 2^{11} \equiv 2048 \pmod{35}$. From here, we can use the division algorithm to divide 2048 by 35, giving $2048 = 58 \cdot 35 + 18$. Thus, $2^{35} \equiv 18 \pmod{35}$.

\newpage

\section{}

Find the last two digits of $123^{403}$.

\subsection{}

We want to find the value of $123^{403} \pmod{100}$. First, we verify that 123 and 100 are coprime, as indeed they are. Now, we find $\phi(100) = \phi(2^2 \cdot 5^2) = \phi(2^2)\phi(5^2) = (4 - 2)(25 - 5) = 2 \cdot 20 = 40$ as per the usual method. Thus, by Euler's theorem, we have that $123^{40} \equiv 1 \pmod{100}$. Using the division algorithm, we have that $403 = 10 \cdot 40 + 3$. As such, $123^{403} \equiv (123^{40})^10 \cdot 123^3 \equiv 1^10 \cdot 123^3 \equiv 123^3 \equiv 1860867 \pmod{100}$. Finally, taking the last two digits of this, we have that $123^3 \equiv 67 \pmod{100}$. Thus, the final two digits of $123^{403}$ are 67.

\newpage

\section{}

Alice chooses primes $p=17$ and $q=23$, as well as public key 7. What is the RSA decryption exponent?

\subsection{}

Here, we start by applying the RSA cryptosystem to generate the modulus $n = pq$, which here is $n = 17 \cdot 23 = 391$. We also compute $\phi(n) = (p - 1)(q - 1) = 16 \cdot 22 = 352$ Next, the public encryption key is $e=7$, and it is trivial to verify that 7 and 352 are coprime. Now, we must find the decryption key $d$ by solving $ed \equiv 1 \pmod{\phi(n)}$, or $7d \equiv 1 \pmod{352}$.

To solve this, we start by knowing that $\gcd(7, 352)$ is of course one, so there will be only one unique solution $d$. Now, we apply the extended Euclidean algorithm to (352, 7).

\begin{align*}
    & \gcd(352,7) & & 352 = 50 \cdot 7 + 2 \\
    = & \gcd(7,2) & & 7 = 3 \cdot 2 + 1 \\
    = & \gcd(2, 1) & & 2 = 2 \cdot 1 + 0 \\
    = & 1
\end{align*}

$$
\begin{matrix}
  & x & y\\
352 & 1 & 0 \\
7 & 0 & 1 \\
2 & 1 & -50 \\
1 & -3 & 151 \\
\end{matrix}
$$

Therefore, the solution to this congruence is $d \equiv 151 \cdot 1 \equiv 151 \pmod{352}$, and it is easy to verify that $7 \cdot 151 \equiv 1057 \equiv 1 \pmod{352}$. Thus, the decryption exponent is $d=151$.

\newpage

\end{document}